Clustering real data with 100 customers (no missing value)

### 2021-07-16 main.Rmd
adding results from new and old method

The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_s(v)$ denotes the probability distribution from observation $s$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points. In our work, we are using $s$ as denoting a household and the underlying variable as the electricity demand. So $f_s(v)$ is the distribution of household $s$ and $v$ is electricity demand.
Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$. Each household consists of a $J*K$ distributions one harmony. We compute the distributional difference
between $(A, B)$ for the $s^{th}$ household using $wpd_{{s}}(A,B)$.
$wpd_{{s}}(A,B)$ denotes the normalized weighted-pairwise distributional distances between $(A, B)$ and is a feature which measures distributional difference between harmonies. If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony. We aim to have pockets of households showing similar periodic behavior by conidering $wpd$ vlaues for different harmonies and some time series features. The features should also characterize probability distributions of different household.

>>>>

The existing work on clustering probability distributions assumes we have an iid sample $f_1(v),\dots,f_n(v)$, where $f_i(v)$ denotes the probability distribution from observation $i$ over some random variable $v = \{v_t: t = 0, 1, 2, \dots, T-1\}$ observed across $T$ time points. In our work, we are using $i$ as denoting a household and the underlying variable as the electricity demand. So $f_i(v)$ is the distribution of household $i$ and $v$ is electricity demand.

We want to cluster distributions of the form $f_{i,j,k}(v)$, where $i$ and $j$ denote

Consider a harmony table consisting of many harmonies, each of the form $(A, B)$, such that $A = \{ a_j: j = 1, 2, \dots, J\}$ and $B = \{ b_k: k = 1, 2, \dots, K\}$. Each household consists of a $J*K$ distributions one harmony. We compute the distributional difference
between $(A, B)$ for the $s^{th}$ household using $wpd_{{s}}(A,B)$.
$wpd_{{s}}(A,B)$ denotes the normalized weighted-pairwise distributional distances between $(A, B)$ and is a feature which measures distributional difference between harmonies. If we have $H_{N_C}$ harmonies in the harmony table, then for each household we have a vector of $wpd_{{s}}$ of $H_{N_C}$ elements with each element corresponding to one harmony. We aim to have pockets of households showing similar periodic behavior by conidering $wpd$ vlaues for different harmonies and some time series features. The features should also characterize probability distributions of different household.

### 2021-07-16 old/realdata-cust10.Rmd
adding results from new and old method

---
title: "Clustering real data with ten customers and choosing optimal number of clusters"
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
    toc: false    
---

@@

```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for 5 households (facets). The data is squeezed stopping us from seeing any behavioral patterns."}
library(lubridate)

@@

```{r hod-box, fig.cap="The raw plots for 5 households shown for Sept 2013-Oct 2013. The data is zoomed in to emphasize weekly, hourly or any behavior/patterns that they might have.", eval = FALSE}
sm %>% 
  create_gran("hour_day") %>% 
  ggplot(aes(x = hour_day, y = general_supply_kwh)) + 
  geom_boxplot() + 
  facet_wrap(~customer_id, ncol = 1, scales = "free_y",
             strip.position = "right") 
```

@@

```{r hod-deciles, fig.cap = "The deciles plots for 5 households are shown for entire observation period across hours of the day. The y-scales are made free to allow us to see daily patterns for each households. The deciles for all households tend to show an increase in the morning and evening hours, although the hours differ. Except for the 90th decile, the deciles for the 4th household look pretty flat."}
data_heatmap %>% 
  ggplot() + geom_line(aes(x=category, 
                           y = quantile_val, 
                           color = quantile_prob)) + facet_wrap(~customer_serial_id, ncol = 1, scales="free_y",
                                                                strip.position = "right") + 
  scale_color_brewer(palette = "Blues") + 
  #scale_y_log10() +
  theme_bw() +
  xlab("hours of the day") +
  ylab("demand (in Kwh)")
```

@@

The following figures show dendograms when we are using optimal number of clusters (k) from k = fpc::nselectboot(), k = 2, 3, 4 and 5. We observe that the $4^{th}$ household tend to get split as different clusters as increase the number of clusters.

>>>>

---
title: "Clustering real data with 10 customers and choosing optimal number of clusters"
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
    toc: false    
---

@@

```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for ten households (facets). The data is squeezed stopping us from seeing any behavioral patterns."}
library(lubridate)

@@

```{r hod-box, fig.cap="The raw plots for ten households shown for Sept 2013-Oct 2013. The data is zoomed in to emphasize weekly, hourly or any behavior/patterns that they might have.", eval = FALSE}
sm %>% 
  create_gran("hour_day") %>% 
  ggplot(aes(x = hour_day, y = general_supply_kwh)) + 
  geom_boxplot() + 
  facet_wrap(~customer_id, ncol = 1, scales = "free_y",
             strip.position = "right") 
```

@@

```{r hod-deciles, fig.cap = "The deciles plots for ten households are shown for entire observation period across hours of the day. The y-scales are made free to allow us to see daily patterns for each households. The deciles for all households tend to show an increase in the morning and evening hours, although the hours differ. Except for the 90th decile, the deciles for the 4th household look pretty flat."}
data_heatmap %>% 
  ggplot() + geom_line(aes(x=category, 
                           y = quantile_val, 
                           color = quantile_prob)) + facet_wrap(~customer_serial_id, ncol = 1, scales="free_y",
                                                                strip.position = "right") + 
  scale_color_brewer(palette = "Blues") + 
  #scale_y_log10() +
  theme_bw() +
  xlab("hours of the day") +
  ylab("demand (in Kwh)")
```

@@

The following figures show dendograms when we are using optimal number of clusters (k) from k = fpc::nselectboot(). We observe that the $4^{th}$ household tend to get split as different clusters as increase the number of clusters.

### 2021-07-16 realdata-cust10-noiter-scaled.Rmd
adding results from new and old method

```{r}
sm <- gravitas::smart_meter10 
```

```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for 10 households (facets). The data is squeezed stopping us from seeing any behavioral patterns."}
library(lubridate)

sm %>% 
  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
  geom_line() +
  facet_wrap(~customer_id, ncol = 1, 
             strip.position = "right") +
  theme_bw() +
  xlab("time") +
  ylab("demand (in Kwh)")
```

```{r raw-sub, fig.cap="The raw plots for ten customers shown for Sept 2013-Oct 2013. The data is zoomed in and the y-scales made free to emphasize weekly, hourly or any behavior/patterns that they might have. There is some daily and weekly pattern in most households."}
sm %>% 
  filter(year(reading_datetime) == 2013 &
           month(reading_datetime) %in% c(9, 10)) %>% 
  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
  geom_line() +
  facet_wrap(~customer_id, ncol = 1, scales = "free_y", 
             strip.position = "right") +
  theme_bw()+
  ylab("demand (in Kwh)") +
  theme(panel.grid.major.x = element_blank()) +
  scale_x_datetime("Date", date_labels = "%b %d",
                   breaks = "1 week",
                   date_minor_breaks = "1 day")  + 
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_line(colour = "#D3D3D3"))

```{r hod}
sm_hod <- gravitas::create_gran(sm, "hour_day") %>% 
  mutate(customer_id = as.integer(customer_id))
```

@@

```{r hod-deciles, fig.cap = "The deciles plots for 10 households are shown for entire observation period across hours of the day. The y-scales are made free to allow us to see daily patterns for each households. The deciles for all households tend to show an increase in the morning and evening hours, although the hours differ. Except for the 90th decile, the deciles for the 4th household look pretty flat."}
data_heatmap %>% 
  ggplot() + geom_line(aes(x=category, 
                           y = quantile_val, 
                           color = quantile_prob)) + 
  facet_wrap(~customer_serial_id, 
             ncol = 1,
             scales="free_y") + 
  scale_color_brewer(palette = "Blues") + 
  #scale_y_log10() +
  theme_bw() +
  xlab("hours of the day") +
  ylab("demand (in Kwh)") +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))
```

@@

```{r hod-dist}
sm_dist_data <- sm_hod_quantiles %>% 
  mutate(customer_serial_id = as.numeric(customer_serial_id)) %>% 
  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
  select(-quantile_prob) %>% 
  group_by(customer_serial_id, categories_serial_id) %>%
  nest() %>% 
  pivot_wider(names_from = categories_serial_id, values_from = data)
```

@@

```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 7, border = 2:8)
```

@@

>>>>

```{r}

elec_ts <- read_rds("../data/elec_ts.rds")

set.seed(12345)

sm_100 <- elec_ts %>%
  as_tibble() %>% 
  distinct(customer_id) %>%
  slice_sample(n = 100)

sm <- elec_ts %>% 
  filter(customer_id %in% sm_100$customer_id) %>% 
  ungroup()

sm_scaled <- sm %>%
  group_by(customer_id) %>% 
  dplyr::mutate(general_supply_kwh = scale(general_supply_kwh)) %>% 
  ungroup()
```

```{r hod}
sm_hod <- gravitas::create_gran(sm_scaled, "hour_day") %>% 
  mutate(customer_id = as.integer(customer_id))
```

@@

```{r hod-deciles, fig.cap = "The deciles plots for 10 households are shown for entire observation period across hours of the day. The y-scales are made free to allow us to see daily patterns for each households. The deciles for all households follow different shape as a result of different customer behavior."}
data_heatmap %>% 
  ggplot() + geom_line(aes(x=category, 
                           y = quantile_val, 
                           color = quantile_prob)) + 
  facet_wrap(~customer_serial_id, 
             ncol = 1,
             scales="free_y") + 
  scale_color_brewer(palette = "Blues") + 
  #scale_y_log10() +
  theme_bw() +
  xlab("hours of the day") +
  ylab("demand (in Kwh)") +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))
```

@@

```{r hod-dist}
sm_dist_data <- sm_hod_quantiles %>% 
  mutate(customer_serial_id = as.numeric(customer_serial_id)) %>% 
  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
  select(-quantile_prob) %>% 
  group_by(customer_serial_id, categories_serial_id) %>%
  nest() %>% 
  pivot_wider(names_from = categories_serial_id, values_from = data)

sm_hod_quantiles %>% 
  filter()

@@

```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 5, border = 2:8)
```

@@

sm_hod <- gravitas::create_gran(sm, "hour_day") %>% 
  mutate(customer_id = as.integer(customer_id))

sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

### 2021-07-16 supplementary/algo2-elec8.Rmd
adding results from new and old method

---
title: "algo2-elec16"
author: "Sayani Gupta"
date: "15/07/2021"
output: pdf_document
---

@@

##----harmony-all
elec_harmony_8local <- read_rds("../paper-hakear/paper/data/elec_harmony_all.rds")

elec_harmony_new8 <- 
read_rds("data/elec_harmony_100-gap.rds") %>% 
  bind_rows(.id = "id") %>% 
  mutate(id = as.numeric(id)) %>% 
  mutate(id = 8 + id) %>% 
  mutate(facet_variable = case_when(
    facet_variable == "hour_day" ~ "hod" ,
    facet_variable == "day_month" ~ "dom" ,
    facet_variable == "day_week" ~ "dow" ,
    facet_variable == "week_month" ~ "wom" ,
    facet_variable == "wknd_wday" ~ "wdwnd"
  )) %>% 
  mutate(x_variable = case_when(
    x_variable == "hour_day" ~ "hod" ,
    x_variable == "day_month" ~ "dom" ,
    x_variable == "day_week" ~ "dow" ,
    x_variable == "week_month" ~ "wom" ,
    x_variable == "wknd_wday" ~ "wdwnd"
  )) %>% 
  mutate(id = paste("id", id, sep = " ")) %>% 
  group_by(id) %>% 
  mutate(rank = row_number())

elec_harmony_all <- bind_rows(elec_harmony_8local, elec_harmony_new8)

@@

data_comb_filter = elec_sig_split %>% 
mutate(low_proxy = if_else(significant=="low", 0, 1))%>% 
select(comb, id, low_proxy) %>% 
pivot_wider(names_from = id, values_from = low_proxy) %>% 
mutate(score = rowMeans(across(`id 1`:`id 16`))) %>% 
filter(score !=0)

@@

hc = stats::hclust(dist(data_clust[-1]),method="complete")
plot(hc)

@@

plot(hc, cex = 0.6)
rect.hclust(hc, k = koptimal, border = 2:10)

groups <- cutree(hc, k=8) %>% as_tibble() %>% 
mutate(id = data_clust$id)

@@

parcoord <- GGally::ggparcoord(data_pcp,
                   columns = 2:ncol(data_pcp),
                   groupColumn = 1,
                   showPoints = TRUE, 
                   title = "(c)",
                   alphaLines = 0.8,
                   scale = "globalminmax"
) + ggplot2::theme_bw() +
  scale_color_viridis(discrete=TRUE) + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")
parcoord

@@

elec_8local <- read_rds(here("../paper-hakear/paper/data/elec_all-8.rds")) %>% 
  dplyr::filter(date(reading_datetime) >= ymd("20190701"), date(reading_datetime) < ymd("20191231"), meter_id==1) %>% 
  select(-meter_id) %>% 
  rename("id" = "household_id",
         "date_time" = "reading_datetime") %>% 
  mutate(date = date(date_time)) %>% 
  mutate(id = paste("id", id, sep = " ")) %>% 
  left_join(groups)%>% 
             rename ("group" = "value")

elec_ts <- read_rds("data/elec_ts.rds")

set.seed(12345)

sm_100 <- elec_ts %>%
  as_tibble() %>% 
  distinct(customer_id) %>%
  slice_sample(n = 100) %>% 
  head(8)

elec_sgsc <- filter(elec_ts, customer_id %in% sm_100$customer_id) %>% 
  ungroup()

sm <- bind_rows(elec_8local,
                elec_sgsc)

sm_hod_dow <- sm %>% 
gravitas::create_gran("hour_day") %>% 
  gravitas::create_gran("day_week")

@@

  sm_hod_dow %>% 
  filter(group==1) %>% 
  gravitas::prob_plot(
     "day_week",
     "hour_day",
  response = "kwh",
  plot_type = "quantile", nrow=1)

  sm_hod_dow %>% 
  filter(group==2) %>% 
  gravitas::prob_plot(  "day_week",
     "hour_day",
  response = "kwh",
  plot_type = "quantile", nrow=1)

  sm_hod_dow %>% 
  filter(group==3) %>% 
  gravitas::prob_plot(  "day_week",
     "hour_day",
  response = "kwh",
  plot_type = "quantile", nrow=1)

  sm_hod_dow %>% 
  filter(group==4) %>% 
  gravitas::prob_plot(  "day_week",
     "hour_day",
  response = "kwh",
  plot_type = "quantile", nrow=1) 
```
<!-- ncol_sm <- seq_len(ncol(sm_hod_list))[-ncol(sm_hod_list)] -->
<!-- nrow_sm <- seq_len(nrow(sm_hod_list)) -->

>>>>

---
title: "Algorithm using wpd values for clustering 8 households"
author: "Sayani Gupta"
date: "13/07/2021"
output: pdf_document
---

@@

##----harmony-all
elec_harmony_all <- read_rds("../../paper-hakear/paper/data/elec_harmony_all.rds")

@@

data_comb_filter = elec_sig_split %>% 
mutate(low_proxy = if_else(significant=="low", 0, 1))%>% 
select(comb, id, low_proxy) %>% 
pivot_wider(names_from = id, values_from = low_proxy) %>% 
mutate(score = rowMeans(across(`id 1`:`id 8`))) %>% 
filter(score !=0)

@@

hc = stats::hclust(dist(data_clust[-1]),method="complete")
#plot(hc)

@@

groups <- cutree(hc, k=4) %>% as_tibble() %>% 
mutate(id = data_clust$id)

@@

parcoord <- GGally::ggparcoord(data_pcp,
                   columns = 2:ncol(data_pcp),
                   groupColumn = 1,
                   showPoints = TRUE, 
                   title = "(c)",
                   alphaLines = 0.8,
                   scale = "globalminmax"
) + ggplot2::theme_bw() +
  scale_color_viridis(discrete=TRUE) + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.y = ggplot2::element_text(angle = 0)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")

@@

elec <- read_rds(here("../paper-hakear/paper/data/elec_all-8.rds")) %>% 
  dplyr::filter(date(reading_datetime) >= ymd("20190701"), date(reading_datetime) < ymd("20191231"), meter_id==1) %>% 
  select(-meter_id) %>% 
  rename("id" = "household_id",
         "date_time" = "reading_datetime") %>% 
  mutate(date = date(date_time)) %>% 
  mutate(id = paste("id", id, sep = " ")) %>% 
  left_join(groups)%>% 
             rename ("group" = "value")

elec_zoom <-  elec %>%
  as_tibble() %>% 
  dplyr::filter(date(date_time) > as.Date("2019-08-01") & date(date_time) < (as.Date("2019-09-30"))) %>%
  ggplot(aes(x=date_time, y = kwh)) +
  #geom_point(size = 0.1, colour = "black", alpha = 0.3) +
  geom_line(size = 0.1, colour = "blue") +
  facet_wrap(~id, 
             scales = "free_y",
             ncol = 1,
             strip.position =  "right") + 
  xlab("Time [30m]") + 
  theme_grey() + 
  ylab("Energy demand (in Kwh)") + ggtitle("(b)") +
  theme(panel.grid.major.x = element_blank()) +
  scale_x_datetime("Date", date_labels = "%b %d",
                   breaks = "1 week",
                   date_minor_breaks = "1 day")  + theme_bw() +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_line(colour = "#D3D3D3")) +
  theme(
    strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))

plot(hc, cex = 0.6)
rect.hclust(hc, k = 4, border = 2:7)

elec_zoom
parcoord

sm_hod_dow <- elec %>% 
gravitas::create_gran("hour_day") %>% 
  gravitas::create_gran("day_week")

@@

  sm_hod_dow %>% 
  filter(group==1) %>% 
  gravitas::prob_plot(
     "day_week",
     "hour_day",
  response = "kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10() +
    ggtitle("Group 1")

  sm_hod_dow %>% 
  filter(group==2) %>% 
  gravitas::prob_plot(  "day_week",
     "hour_day",
  response = "kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10()+
    ggtitle("Group 2")

  sm_hod_dow %>% 
  filter(group==3) %>% 
  gravitas::prob_plot(  "day_week",
     "hour_day",
  response = "kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10() +
    ggtitle("Group 3")

  sm_hod_dow %>% 
  filter(group==4) %>% 
  gravitas::prob_plot("day_week",
     "hour_day",
  response = "kwh",
  plot_type = "quantile", nrow=1) +
  scale_y_log10() +
    ggtitle("Group 4")
```
<!-- ncol_sm <- seq_len(ncol(sm_hod_list))[-ncol(sm_hod_list)] -->
<!-- nrow_sm <- seq_len(nrow(sm_hod_list)) -->

### 2021-07-16 supplementary/realdata-cust100-noiter-scaled.Rmd
adding results from new and old method

```{r hod-deciles, fig.cap = "The deciles plots for 10 households are shown for entire observation period across hours of the day. The y-scales are made free to allow us to see daily patterns for each households. The deciles for all households follow different shape as a result of different customer behavior."}
data_heatmap %>% 
  ggplot() + geom_line(aes(x=category, 
                           y = quantile_val, 
                           color = quantile_prob)) + 
  facet_wrap(~customer_serial_id, 
             ncol = 1,
             scales="free_y") + 
  scale_color_brewer(palette = "Blues") + 
  #scale_y_log10() +
  theme_bw() +
  xlab("hours of the day") +
  ylab("demand (in Kwh)") +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))
```

@@

```{r hod-dist}
sm_dist_data <- sm_hod_quantiles %>% 
  mutate(customer_serial_id = as.numeric(customer_serial_id)) %>% 
  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
  select(-quantile_prob) %>% 
  group_by(customer_serial_id, categories_serial_id) %>%
  nest() %>% 
  pivot_wider(names_from = categories_serial_id, values_from = data)

sm_hod_quantiles %>% 
  filter()

>>>>

@@

```{r hod-dist}

sm_filter_nozerosum <- sm_hod_quantiles %>% 
  pivot_longer(-c(1, 2), names_to = "quantiles", values_to = "quantvalue") %>% 
  group_by(customer_serial_id) %>% 
  summarise(m = sum(quantvalue)) %>% 
  filter(m!=0)

sm_hod_quantiles <- sm_hod_quantiles %>%
  filter(customer_serial_id %in% sm_filter_nozerosum$customer_serial_id)

sm_dist_data <- sm_hod_quantiles %>% 
  mutate(customer_serial_id = as.numeric(customer_serial_id)) %>% 
  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
  select(-quantile_prob) %>% 
  group_by(customer_serial_id, categories_serial_id) %>%
  nest() %>% 
  pivot_wider(names_from = categories_serial_id, values_from = data)

### 2021-07-16 supplementary/realdata-cust100-noiter.Rmd
adding results from new and old method

---
title: "Clustering real data with ten customers without iterations"
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
    toc: false    
---

@@

# Plots of raw data

Five households are considered for the following analysis from the SGSC data set.
The data sets contain three columns `customer_id`, `reading_datetime` and `general_supply_kwh`. They consist of half-hourly data from 2012-2014. Figure \ref{fig:raw-all} shows how the raw time plots for these five households look like. Since all the dataset looks squeezed on this linear time scale, Figure \ref{fig:raw-sub} shows how the raw plot looks for 2 months (Sept 2013 and Oct 2013).

```{r}

elec_ts <- read_rds("../data/elec_ts.rds")

set.seed(12345)

sm_100 <- elec_ts %>%
  as_tibble() %>% 
  distinct(customer_id) %>%
  slice_sample(n = 100)

sm <- elec_ts %>% 
  filter(customer_id %in% sm_100$customer_id) %>% 
  ungroup()

sm_scaled <- sm %>%
  group_by(customer_id) %>% 
  dplyr::mutate(general_supply_kwh = scale(general_supply_kwh)) %>% 
  ungroup()
```

```{r hod}
sm_hod <- gravitas::create_gran(sm_scaled, "hour_day") %>% 
  mutate(customer_id = as.integer(customer_id))
```

@@

```{r}
quantile_prob_val = seq(0.1, 0.9, 0.1)

@@

```{r hod-dist}

@@

sm_dist_data <- sm_hod_quantiles %>% 
  mutate(customer_serial_id = as.numeric(customer_serial_id)) %>% 
  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
  select(-quantile_prob) %>% 
  group_by(customer_serial_id, categories_serial_id) %>%
  nest() %>% 
  pivot_wider(names_from = categories_serial_id, values_from = data)

@@

@@

@@

The following figures show dendograms when we are using optimal number of clusters (k) from k = fpc::nselectboot(), k = 2, 3, 4 and 5. We observe that the $4^{th}$ household tend to get split as different clusters as increase the number of clusters.

@@

library(fpc)
koptimal = fpc::nselectboot(as.dist(dist_mat_format),
                        B = 50,
                        method = "complete",
                        clustermethod = disthclustCBI, 
                        classification = "averagedist",
                        krange = 2:10)

@@

mds_plot_10 <- ggplot(all_data_cluster,
       aes(x = Dim.1,
           y = Dim.2,
       color = groups,
       label = rownames(mds))) +
  geom_point(size = 1) +
  geom_text(check_overlap = TRUE)  +
  theme_classic()+
  theme(legend.position = "bottom") +
  scale_color_brewer(palette = "Dark2")

@@

_k = 7_

```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 5, border = 2:8)
```

@@

```{r hod-heatmap, fig.cap="Heatmaps show deciles on the y-axis with customers on the x-axis with colors filled by the value of deciles and faceted by hours of the day. It again seems like all the households have gradual change in colors (almost for all hours) as we move to higher deciles except for the 4th household. "}
quantile_prob_val = c(0.25, 0.5, 0.75)

sm_hod <- gravitas::create_gran(sm, "hour_day") %>% 
  mutate(customer_id = as.integer(customer_id))

sm_hod_list <- sm_hod %>% 
  as_tibble() %>% 
  select(customer_id, hour_day, general_supply_kwh) %>% 
  pivot_wider(names_from = hour_day,
              values_from = general_supply_kwh)

groups_ref <- groups %>% as_tibble %>% 
              mutate(customer_serial_id = sm_hod_list$customer_id) %>% 
             rename ("group" = "value")

sm_hod_list_cat <-  sm_hod_list %>% left_join(groups_ref, by = c("customer_id" = "customer_serial_id")) %>% select(group, everything())

ncol_sm <- seq_len(ncol(sm_hod_list_cat[-c(1, 2)]))
nrow_sm <- unique(sm_hod_list_cat$group)

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list_cat%>% 
     dplyr::filter(group == x) %>% 
     select(-c(1, 2)) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_val)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "group_id") %>% 
  mutate(group_id = as.integer(group_id),
         categories_serial_id = as.integer(categories_serial_id))

categories_ref_cat <- names(sm_hod_list_cat)[-c(1, 2)] %>% as_tibble() %>% 
  mutate(categories_serial_id = row_number())

@@

# data_heatmap%>% 
# ggplot(aes(x= as.factor(group_id), y = as.factor(quantile_prob))) +
#   geom_tile(aes(fill = quantile_val)) +
#   scale_fill_distiller(palette = "YlGnBu")+ 
#   facet_wrap(~category) +
#   theme(axis.text.x = element_text(angle = 90)) +
#   ylab("deciles") +
#   xlab("customers")
```

```{r cluster-characterization}
data_heatmap %>% 
  pivot_wider( names_from = quantile_prob, 
               values_from = quantile_val) %>% 
  ggplot(aes(x = category)) +  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "lightblue") +
  geom_line(aes(y = `50%`)) +
  facet_wrap(~group_id, ncol = 1, scales = "free_y", labeller = "label_both") +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))
```

>>>>

---
title: "Clustering real data with 100 customers (no missing value)"
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
    toc: false
    cache: TRUE
---

@@

```{r read-data}
# elec_ts <- read_rds("../data/elec_ts.rds")
# 
# set.seed(12345)
# 
# sm_100 <- elec_ts %>%
#   as_tibble() %>% 
#   distinct(customer_id) %>%
#   slice_sample(n = 100)
# 
# sm <- elec_ts %>% 
#   filter(customer_id %in% sm_100$customer_id) %>% 
#   ungroup()

#elec_nogap <- read_rds("../data/elec_nogap.rds")
```

```{r read100}
# 
# set.seed(12345)
# 
# sm_100 <- elec_nogap %>%
#   as_tibble() %>%
#   distinct(customer_id) %>%
#   slice_sample(n = 100)
# 
# sm <- elec_nogap %>%
#   filter(customer_id %in% sm_100$customer_id) %>%
#   ungroup()
# write_rds(sm, "../data/elec_nogap-100cust.rds")

sm <- read_rds("../data/elec_nogap-100cust.rds")
```

```{r hod}
sm_hod <- gravitas::create_gran(sm, "hour_day") 
```

@@

```{r heatmpa-data}
quantile_prob_val = seq(0.1, 0.9, 0.1)

@@

@@

@@

```{r hod-dist}
sm_dist_data <- sm_hod_quantiles %>% 
  mutate(customer_serial_id = as.numeric(customer_serial_id)) %>% 
  pivot_longer(-c(1, 2), names_to = "quantile_prob", values_to = "quantile_val") %>% 
  select(-quantile_prob) %>% 
  group_by(customer_serial_id, 
           categories_serial_id) %>%
  nest() %>% 
  pivot_wider(names_from = categories_serial_id, 
              values_from = data) %>% 
  ungroup()
```

@@

lookup_tbl <- sm_hod_list %>% 
  mutate(customer_serial_id = row_number()) %>% 
  select(customer_id, customer_serial_id) %>% 
  filter(customer_serial_id %in% sm_filter_nozerosum$customer_serial_id)

@@

@@

library(fpc)
koptimal = fpc::nselectboot(as.dist(dist_mat_format),
                        B = 500,
                        method = "complete",
                        clustermethod = disthclustCBI, 
                        classification = "averagedist",
                        krange = 2:15)

@@

mds_plot_10 <- ggplot(all_data_cluster,
       aes(x = Dim.1,
           y = Dim.2,
       color = groups,
       label = rownames(mds))) +
  geom_point(size = 1) +
  geom_text(check_overlap = TRUE)  +
  theme_classic()+
  theme(legend.position = "bottom")

@@

_k = 15_

```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 15, border = 2:20)
```

@@

```{r hod-heatmap, fig.cap="Heatmaps show deciles on the y-axis with customers on the x-axis with colors filled by the value of deciles and faceted by hours of the day. It again seems like all the households have gradual change in colors (almost for all hours) as we move to higher deciles except for the 4th household. "}
quantile_prob_val = c(0.1, 0.25, 0.5, 0.75, 0.9)

groups_ref <- groups %>% as_tibble %>% 
              mutate(customer_id = lookup_tbl$customer_id,
                     customer_serial_id = lookup_tbl$customer_serial_id) %>% 
             rename ("group" = "value")

groups_g3 <- groups_ref %>% 
  group_by(group) %>% 
  count() %>% 
  filter(n>=3)

sm_hod_list_cat <- sm_hod_list %>% left_join(groups_ref) %>%
  filter(customer_serial_id %in% sm_filter_nozerosum$customer_serial_id) %>% 
select(group, customer_serial_id, everything())

ncol_sm <- seq_len(ncol(sm_hod_list_cat[-c(1, 2, 3)]))
nrow_sm <- unique(sm_hod_list_cat$group)

sm_hod_quantiles_cat <- map(nrow_sm, function(x){
  map(ncol_sm, function(y){
   cell <- sm_hod_list_cat%>% 
     dplyr::filter(group == x) %>% 
     select(-c(1, 2, 3)) %>% 
     extract(y) %>% 
     unlist()
   quantile(cell, prob = quantile_prob_val)
})  %>% bind_rows(.id = "categories_serial_id")
}) %>% bind_rows(.id = "group_id") %>% 
  mutate(group_id = as.integer(group_id),
         categories_serial_id = as.integer(categories_serial_id))

categories_ref_cat <- names(sm_hod_list_cat)[-c(1, 2, 3)] %>% as_tibble() %>% 
  mutate(categories_serial_id = row_number())

@@

```{r cluster-characterization}
data_heatmap %>% 
  filter(group_id %in% groups_g3$group) %>% 
  pivot_wider( names_from = quantile_prob, 
               values_from = quantile_val) %>% 
  ggplot(aes(x = category)) +  
  geom_ribbon(aes(ymin = `25%`, 
                  ymax = `75%`), fill = "#D55E00", alpha = 0.8) +
  geom_ribbon(aes(ymin = `10%`, 
                  ymax = `90%`), fill = "#56B4E9", alpha = 0.4) + 
  geom_line(aes(y = `50%`)) +
  facet_wrap(~group_id, ncol = 2,
             scales = "free_y",
             strip.position = "right") +
  theme_bw() +
    theme(strip.text = element_text(size = 10, margin = margin(b = 0, t = 0)))
```
